---
title: "MATH 216 Homework 4"
author: "Andrea Narciso"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Suggested packages; feel free to use more!
library(tidyverse)
library(USAboundaries)
library(maptools)
library(sp)
library(broom)
library(stringr)


```



## Admistrative:

Please indicate

* Who you collaborated with: 
* Roughly how much time you spent on this HW so far: 5 hours
* The URL of the RPubs published URL [here](http://rpubs.com/anarciso/HW-4).
* What gave you the most trouble: downloading the census data, figuring out how to find the weighted distance, how to use/combine the different data sets
* Any comments you have:



## Question 1:

**Question**:  DHL, a German shipping company, has a ton of money to build a
shipping hub in the United States. Where should they build it?

**Deliverable**: A choropleth map of the lower 48 states (no Puerto Rico) where
the color corresponds to the distance of the given area to the shipping hub,
which is marked on the map with a large red dot.

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Perform your computations here:

# Counties: shape data
counties_shp <- us_counties()
states_data <- counties_shp@data

states_polygon <- tidy(counties_shp, region="geoid")
states <- left_join(states_polygon, states_data, by=c("id"="geoid"))

# Centroids
centroids <- coordinates(counties_shp)
states_centroids <- cbind(centroids, states_data) %>%  #states_data: OG of states
  mutate(geoid = as.integer(geoid)) 

# Counties: census data
USA_census <-
  read_csv("census_data_USA_2010.csv") %>% 
  rename(pop=SE_T001_001) %>% 
  mutate(
    state_code = str_pad(Geo_STATE, 2, pad="0"),
    county_code = str_pad(Geo_COUNTY, 3, pad="0")
  )
USA_census <- unite(USA_census, "FIPS_code", state_code, county_code, sep = "") %>% 
  mutate(FIPS_code = as.integer(FIPS_code)) 

# Combines all and finds weighted long and lat values
census_centroids <- left_join(USA_census, 
                              states_centroids, by=c("FIPS_code"="geoid")) %>%
  filter( !state_name %in% c("Alaska", "Hawaii", "Puerto Rico")) %>% 
  rename(long = `1`, lat = `2`) %>% 
  mutate(weighted_long = long*pop,
         weighted_lat = lat*pop)

combined_data <- left_join(census_centroids, 
                           states_polygon, by=c("Geo_FIPS"="id")) 

# Location of hub
hub_x_coord <- sum(census_centroids$weighted_long, na.rm=TRUE) / sum(census_centroids$pop, na.rm=TRUE)
hub_y_coord <-sum(census_centroids$weighted_lat, na.rm=TRUE) / sum(census_centroids$pop, na.rm=TRUE)

hub_x_coord
hub_y_coord

# Distance from Hub
dist_fr_hub <- combined_data %>% 
  mutate(dist = sqrt((long.y - hub_x_coord)^2 + (lat.y - hub_y_coord)^2)) 

# NEED TO FIX VIRGINIA DATA !!!



```
DHL should put their hub at (-91.64711, 37.55005).

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Here is a template of a ggplot call; feel free to change this to suit your
# needs. Once your computations above are complete, set the eval=TRUE on the
# line above to tell R Markdown to run this code block.

ggplot(data=dist_fr_hub, aes(x=long.y, y=lat.y, group=group, fill=dist)) +
  # Plot choropleth polygons, low -> high being colored white -> dark green
  geom_polygon() +
  scale_fill_gradient(low="white", high="darkgreen") + 
  # Trace outlines of areas
  geom_path(size=0.1) +
  # Mark hub
  annotate("point", x=hub_x_coord, y=hub_y_coord, col="red", size=5) +
  # Correct aspect ratio for map
  coord_map()
```





